Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2021-10-10 methods_paper/fabl-paper.tex
minor edit.

\subsection{Beta Linkage Model}

One criticism of using mixture models is that the decision rule will lead to "many-to-many assignments," breaking one-to-one matching assumptions. One way of resolving this issue to incorporate the constraint directly into the model as follows

%our model can be written as a Bayesian model for comparison data. In general, it has the same form of \cite{sadinle_bayesian_2017} and can be written as follows:
\begin{align}
	&\bm{\Gamma}_{ij} \mid Z_j =i \stackrel{iid}{\sim} \mathcal{M}(\bm{m})  \\
	&\bm{\Gamma}_{ij} \mid Z_j \neq i \stackrel{iid}{\sim} \mathcal{U}(\bm{u}) \notag \\
	& \bm{Z} \sim \mathcal{B}, \notag
\end{align}
where $\mathcal{B}$ represents a prior on the space of bipartite matchings, and $\mathcal{M}(\bm{m}), \mathcal{U}(\bm{u})$ are models for the comparison vectors of matches and non-matches. For \texttt{BRL} \cite{sadinle_bayesian_2017} proposed the \emph{beta distribution for bipartite matching}, given by
$$P(\mathbf{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\mathbf{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\mathbf{Z}) + \alpha_{\pi}, n_2 - \mathbf{Z}) - \beta_{\pi}])}{\text{B}(\alpha_{\pi}, \beta_{\pi})}
$$
where $\text{B}(\cdot, \cdot)$ represents the Beta function, and hyper parameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$. This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate
records when sampling \(Z_j\). This creates a dependency that makes the
sampler \emph{inherently serial}, which makes the sampler slow when working on large linkage tasks.

@@

------------------------------------------------------

\textcolor{black}{I think its best just not to mention probabalistic blocking. My method doesn't use it, and introducing it is really confusing (Sadinle 2017, Enamorado 2019, and Steorts 2014/2015 don't mention it).}

We note here that the word "hashing" is also used in probabalistic blocking, which uses techniques from machine learning to cluster records together before attempting to link records. The use of such probabalistic blocking to provide even greater computational gains will be explored in future work.

-------------------------------------------------------

>>>>

\subsection{Beta Record Linkage Model}

One criticism of using mixture models is that the decision rule will lead to "many-to-many assignments," breaking one-to-one matching assumptions. One way of resolving this issue to incorporate the constraint directly into the model as follows:
%our model can be written as a Bayesian model for comparison data. In general, it has the same form of \cite{sadinle_bayesian_2017} and can be written as follows:
\begin{align}
	&\bm{\Gamma}_{ij} \mid Z_j =i \stackrel{iid}{\sim} \mathcal{M}(\bm{m})  \\
	&\bm{\Gamma}_{ij} \mid Z_j \neq i \stackrel{iid}{\sim} \mathcal{U}(\bm{u}) \notag \\
	& \bm{Z} \sim \mathcal{B}, \notag
\end{align}
where $\mathcal{B}$ represents a prior on the space of bipartite matchings, and $\mathcal{M}(\bm{m}), \mathcal{U}(\bm{u})$ are models for the comparison vectors of matches and non-matches. For \texttt{BRL}, \cite{sadinle_bayesian_2017} proposed the \emph{beta distribution for bipartite matching}, given by
$$P(\mathbf{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\mathbf{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\mathbf{Z}) + \alpha_{\pi}, n_2 - \mathbf{Z}) - \beta_{\pi}])}{\text{B}(\alpha_{\pi}, \beta_{\pi})}
$$
where $\text{B}(\cdot, \cdot)$ represents the Beta function, and hyper parameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$. This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate
records when sampling \(Z_j\). This creates a dependency that makes the
sampler \emph{inherently serial}, which makes the sampler slow when working on large linkage tasks.

@@

%------------------------------------------------------
%
%\textcolor{black}{I think its best just not to mention probabalistic blocking. My method doesn't use it, and introducing it is really confusing (Sadinle 2017, Enamorado 2019, and Steorts 2014/2015 don't mention it).}
%
%We note here that the word "hashing" is also used in probabalistic blocking, which uses techniques from machine learning to cluster records together before attempting to link records. The use of such probabalistic blocking to provide even greater computational gains will be explored in future work. 
%
%-------------------------------------------------------
